---
title: "Topic Modeling"
author: "Jamie Reilly, Ph.D."
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: true
    css: "https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/jr_20.css?raw=true"
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", 'stringi', 'quanteda', 'readr', 'textclean', 'readtext', "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'sotu', 'udpipe', 'caret', 'text2vec', 'textmineR', 'topicmodels',  'quanteda.textmodels', 'seededlda', 'tm', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
```

# TOPIC MODELS
[Link to Kasper Welbers YouTube tutorial about LDA](https://www.youtube.com/watch?v=4YyoMGv1nkc) <br>
<br>

Annotate each of these code snippets as you watch the YT video
```{r}
library(quanteda)

#splits corpus into paragraphs
inaugurations_raw <- corpus_reshape(data_corpus_inaugural, to = "paragraphs")

#tokenize the raw data and remove stopwords from the SMART set
inaugurations_tokenized <- quanteda::tokens(inaugurations_raw, remove_punct = TRUE) %>% tokens_remove(stopwords("SMART"))


#turn tokenized set to document feature matrix
inaugurations_dfm <- dfm(inaugurations_tokenized)

#pull only the tokens that show in at least 5 documents
inaugurations_dfm_freq5 <- dfm_trim(inaugurations_dfm, min_docfreq = 5)
print(inaugurations_dfm_freq5)

library(topicmodels)

#convert to a topicmodels format to be accepted by topicmodels package
inaugurations_dtm_topmod <- convert(inaugurations_dfm_freq5, to = "topicmodels") 

```


# Run the topic model and analyze
Using LDA-based topic model
```{r}

#k is how many clusters/topics you want the function to find
# alpha means how spread out topics per documents are
#higher alpha results in more topics captured per document
set.seed(193)
inaugurations_topicmod  <- LDA(inaugurations_dtm_topmod, method = "Gibbs", k = 5,  control = list(alpha = 0.1))

#object created by LDA
inaugurations_topicmod

#look at top terms per topic
topicmodels::terms(inaugurations_topicmod, 6)

#annotate me
topic = 5

#get the posterior distribution of words and documents which can be used to plot word cloud
#you get the probability that each of those words will be associated with a given topic
words <- topicmodels::posterior(inaugurations_topicmod)$terms[topic, ]

#show highest probability words per topic
topwords <- head(sort(words, decreasing = T), n=50)
head(topwords)

library(wordcloud)

#create wordcloud
wordcloud(names(topwords), topwords)

#look at documents per topic based on posterior probability
#this picks topic 6 specifically
topic.docs = posterior(m)$topics[, topic] 

#sort by the highest probability documents
topic.docs = sort(topic.docs, decreasing=T)
head(topic.docs)

#take the name of first document 
topdoc = names(topic.docs)[1]

#make a corpus from that document
topdoc_corp = corp[docnames(corp) == topdoc]

#annotate me
texts(topdoc_corp)

#match rows from dfm and the names together
docs <- docvars(dfm)[match(rownames(dtm), docnames(dfm)),]

#summarize the statistics into means based on the author of the documents
tpp = aggregate(posterior(m)$topics, by=docs["President"], mean)

#rename the rows
rownames(tpp) = tpp$President

#visualize data into a heat map
heatmap(as.matrix(tpp[-1]))
```


# Code from Jamie
## Load cleaning genie
Load whatever cleaning function you work with in the background
```{r, eval=T}

jamie_genie <- function(dat, wordcol, omit_stops = TRUE, lemmatize = TRUE) {
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(stringi)
  library(textstem)
  
# Convert input object to dataframe
  dat <- as.data.frame(dat)
  
# Input validation, tell R which column in your df is target text, put it in quotes!
  # Check if wordcol is a character string and specifies a valid column
  if (!is.character(wordcol) || length(wordcol) != 1) {
    stop("Oops! You forgot to specify the name of the column in your dataframe (in quotes) that contains your target text for cleaning.")
  }
  
  if (!wordcol %in% names(dat)) {
    stop(paste("Column", wordcol, "not found in dataframe"))
  }

#load stopword and replacement lists from github  
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Temple_stops25.rda?raw=true"))
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/replacements_25.rda?raw=true"))


# Create working copy and perform initial split, standardizes text encoding so all characters are UTF8, adds an id variable
# transform all text to lower
  dat_prep <- dat %>% dplyr::mutate(id_row_orig = factor(seq_len(nrow(dat))),
      text_initialsplit = tryCatch(stringi::stri_enc_toutf8(as.character(.[[wordcol]]),
                                 is_unknown_8bit = TRUE,
                                 validate = TRUE), error = function(e) stringi::stri_encode(as.character(.[[wordcol]]), to = "UTF-8")) %>% tolower())

  # Standardize apostrophes in original text, replace hyphens with spaces
  dat_prep <- dat_prep %>% 
  dplyr::mutate(
    text_initialsplit = ifelse(
      is.na(text_initialsplit), 
      NA_character_, 
      stringi::stri_replace_all_regex(text_initialsplit,
        "[\u2018\u2019\u02BC\u201B\uFF07\u0092\u0091\u0060\u00B4\u2032\u2035]", "'")),
    text_initialsplit = stringi::stri_replace_all_fixed(text_initialsplit, 
      "-", " "))

  # Perform initial split into words (retains contractions)
  dat_prep <- dat_prep %>% tidyr::separate_rows(text_initialsplit, sep = "[[:space:]]+") %>%
    dplyr::mutate(
      text_initialsplit = ifelse(
        is.na(text_initialsplit) | stringi::stri_isempty(text_initialsplit),
        NA_character_,
        text_initialsplit)) %>%
    # Remove original column
    dplyr::select(-all_of(wordcol))

  # Initialize cleaning column with actual cleaned text
  dat_prep <- dat_prep %>%
    dplyr::mutate(
      word_clean = text_initialsplit,
      # Remove non-alphabetic characters except apostrophes
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^a-zA-Z']", " ")
      ),
      # Squish whitespace
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringr::str_squish(word_clean)
      ),
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^a-z']", "")
      ),
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^[:alnum:]']", "")))
  
  
  #Replace words function - this finds contractions and replaces them with their expanded forms from a lookup database. 
 replace_words <- function(word_clean, replacement_df) {
 for (i in seq_len(nrow(replacement_df))) {
   text_vector <- gsub(
     pattern = replacement_df$word[i],
     replacement = replacement_df$replacement[i],
     x = word_clean, ignore.case = TRUE)
 }
 return(word_clean)
 }
}
```

#  Dataset: SOTU Package
The state of the union `sotu` package contains every US Presidential state of the union address
[sotu package](https://cran.r-project.org/web/packages/sotu)

```{r}
library(sotu)
sotu <- cbind(sotu_text, sotu_meta)
sotu$id <- seq_along(sotu$year)
head(sotu)
```

# Model 1: Obama sotu 2016
## clean and lemmatize
use the TM package; data need to be structured as one narrative per row per document. We only have one document (President Obama's transcribed sotu 2016)
```{r, eval=T}
obama16 <- sotu %>% dplyr::filter(id == "236")
colnames(obama16)
obama_clean <- jamie_genie(obama16, wodcol='sotu_text')  #run cleaning genie on obama

```

## Convert to Corpus and DFM
```{r, eval=T}
obama_corpus <- quanteda::corpus(obama_lemma)
obama_dfm <-quanteda::dfm(obama_corpus)
print(obama_dfm)
```

## build model quanteda
lda seeded from `seededlda` package. Let's try a 5 topic model (k=5).
```{r, eval=T}
obama_mod <- textmodel_lda(obama_dfm, k=5)
terms(obama_mod, 15)  #prints the 15 most frequent terms associated with each topic
```

# Model 2: All sotus after 1950
Isolate all SOTU speeches after 1950. These are contained in a dataframe with each sotu speech as a character vector along with metadata in other columns
```{r}
modern <- both %>% filter(year > 1950) %>% mutate(my_text = clean(sotu_text)) 
recent <- modern %>% select(2,3,5,7,8)
head(recent)
#test <- aggregate(sotu_text ~ id, data=modern, FUN=paste, collapse= " ")
```

## read into quanteda maintaining metadata
retain 'president' and 'year' data,  document id is "id" 
```{r}
mydocvars <- c("party", "president", "year")
sotu_corp <- quanteda::corpus(recent, docid_field = "id", text_field = "my_text", docvars=mydocvars)
sotu_tok <- quanteda::tokens(sotu_corp)
sotu_dfm <-  quanteda::dfm(sotu_corp)
topfeatures(sotu_dfm, 5)
textplot_wordcloud(sotu_dfm)
```

# plot lexical dispersion
quanteda calls this an x-ray plot. It outputs a ggplot2 object, so you can accordingly later ggplot geoms
```{r, fig.height=11}
head(docvars(sotu_tok))
tplot1 <- textplot_xray(kwic(sotu_tok, pattern = "war"), kwic(sotu_tok, pattern = "taxes"), sort=TRUE) 
tplot1 + aes(color = keyword) + scale_color_manual(values = c('red', 'blue'))
textplot_xray(kwic(sotu_tok, pattern = "patriots"), kwic(sotu_tok, pattern = "unity"), sort=TRUE)
```

## plot top frequency
```{r, fig.width=9, fig.height=7}
feats <- textstat_frequency(sotu_dfm, n = 50)
# Sort by reverse frequency order
feats$feature <- with(feats, reorder(feature, -frequency))
ggplot(feats, aes(x = feature, y = frequency)) +
    geom_point(color="red") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## plot adjacency
from the topicmodels package - builds a CTM object (correlated term matrix)
```{r}
mod1 <- CTM(sotu_dfm, k=8, method = "VEM", control = NULL)
build_graph(mod1, 0, and = TRUE)  #lambda I set to 0 here - can be 1, don't know why!
```

## feature co-occurence network
```{r, fig.height=8, fig.width=10}
sotu_trim<- dfm_trim(sotu_dfm, min_termfreq = 20)  #trim dfm by frequency
topfeatures(sotu_trim)
nfeat(sotu_trim)
sotu_fcm <- fcm(sotu_trim)
dim(sotu_fcm)
sotu_top_names <- names(topfeatures(sotu_fcm, 50))
sotu_top50 <- fcm_select(sotu_fcm, pattern = sotu_top_names, selection = "keep")
dim(sotu_top50)
size <- log(colSums(dfm_select(sotu_dfm, sotu_top_names, selection = "keep")))
set.seed(123)
textplot_network(sotu_top50, min_freq = 0.8, vertex_size = size / max(size) * 3)
```

## wordfish document similarity
```{r, eval=T, fig.height=11, fig.width=9}
sotu_fish <- textmodel_wordfish(sotu_dfm, dir = c(6, 5))
summary(sotu_fish)
textplot_scale1d(sotu_fish, groups = sotu_dfm$party)
```

## Topic Models (LDA fit)
Build an 8-topic model
```{r, eval=T}
obama_topics <- LDA(obama_dfm, k = 8, control = list(seed = 123))  #let's try an 8-topic model
topic_probs <- tidy(obama_topics, matrix = "beta")  
head(topic_probs, n=20)
```

# Evaluating topic model fit
There could be one topic in a document (e.g., "ice is cold") or many topics.  One challenge is that the number of topics (parameter `k`) is user-specified. We must then evaluate the fit of specific topic models. For a nice summary of some of the fit parameters and guts of Latent Dirichlet Allocation (LDA) visit [Joe Cristian's Rpubs website](https://rpubs.com/jojoecp/643113). <br />

Here are some of the specific parameters we need to consider: <br />
1. `Theta` $\Theta$ represents topic distributions where $\Theta{m}$ is the topic distribution for document m <br />
2. `Phi` $\Phi$ represents <br />
3. `Beta` $\beta$ represents per topic word probabilities with lower betas indicating a smaller mixture of words per topic. <br />
4. `k` number of topics (user specified) <br />
5. `M` is the number of documents <br />
6. `N` is the number of words <br />

The `textmineR` package provides numerous fit measures. We will need to build a topic model from this package working from a document feature matrix. Let's try it out on the sotu_dfm we created earlier. We can use gibbs sampling or VEM variational expectation maximization. We'll use the `FindTopicsNumber` function from the `ldatuning` package to take a shot at determining the best number of topics from 10-70 in all the SOTU addresses from 1950 onward.

```{r}
result <- FindTopicsNumber(
  sotu_dfm,
  topics = seq(from = 10, to = 70, by = 1),
  metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
  method = "Gibbs",
  control = list(seed = 77),
  mc.cores = 2L,
  verbose = TRUE
)
summary(result)
FindTopicsNumber_plot(result)
```



Code from [Kasper Welbers](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/r_text_lda.md)
