---
title: "NLP CHEATSHEET"
author: "Shawn Wang"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    css: !expr here::here("./stylesTemplates/sw_1.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", 
           "RCurl", "psych", "knitr", 
           "RColorBrewer", 'dplyr', "tidytext",
           "tm", "stringi", "textstem",
           install=T)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'gutenbergr', 'janeaustenr', 'quanteda', 'readtext', install=T)
library("quanteda.textplots")
library(quanteda.textstats)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)

library(tidytext)

```


# FUNCTIONS



## Functions to add numbers
```{r}
#function to add 2 numbers

add_func <- function(x, y) {
  
  #add them together
  sum <- x + y
  return(sum)
}



```

<br>

## Function to mean 2 numbers
```{r}


mean_func <- function(x, y) {
  
  #take the sum
  sum <- x + y
  
  #divide by 2
  mean <- sum / 2
  return(mean)
}


```

<br>

## Function to remove any word from a string
```{r}

#old_Word is the word you want to replace
#replacement_Word is the word you want to replace it with
#takes a string returns a string
replace_func <- function(string, old_Word, replacement_Word) {
  
  x <- string
  x_clean <- x %>%
    gsub(pattern = old_Word, replacement = replacement_Word) 
  
  return(x_clean)
  
  
}


```
<br>

## Function to remove special characters
```{r}

#takes a string returns a string
remove_special_func <- function(myString) {
  

  #replaces punctuation and special characters with blanks
  x_clean <- gsub("[[:punct:]]", " ", myString)

  
  return(x_clean)
  
  
}


```
<br>

## Function to split string into data frame, one word per row
```{r}

#takes a string returns a data frame
split_string <- function(x) {
  
  #convert to a data frame first
  cleaned_x <- data.frame(string = x) %>%
    separate_longer_delim(string, delim = " ") %>% #pivots to long format based on delimeter of space
    filter(string != "") #filters so that empty rows are deleted
  return(cleaned_x)
  
}



```

<br>

## Function to remove "the"
```{r}


remove_the <- function(x) {
  
  #replaces "the" with blank
  x_clean <- x %>%
    gsub(pattern = "the", replacement = "", ignore.case = TRUE)
  return(x_clean)
}

```


<br>

## Remove numbers
```{r}


#takes a string returns a string
remove_numbers <- function(x) {
  
  #uses regex to remove any numbers from a string
  x_clean <- x %>%
    gsub(pattern = "\\d+", replacement = " ")
  return(x_clean)
  
}

```

<br>

## Remove non-alphabetic characters, numbers, punctuation
```{r}
keep_alphab <- function(x) {
  
  #uses regex to remove any numbers from a string
  x_clean <- x %>%
    mutate(string = gsub(pattern = "[^a-zA-Z]", replacement = " ", string))
    
    
  
  return(x_clean)
  
}
```



<br>

## Function to clean built from other functions

```{r}

#takes a string, returns a tidy data frame with one word per row
#x is the data frame
#words is a list of words that we want to remove as a vector
clean_string <- function(x, words) {
  

  
  #keep only alphabetic characters
  #remove non-alphabetic characters, numbers, punctuation
  x <- keep_alphab(x)
  
  
  
  #remove words that we don't want by sequencing along a list of words that we don't want in a loop
  for (i in seq_along(words)) {
    
    x <- replace_func(string = x, old_Word = words[i], replacement_Word = "")
    
  }
  
  x <- gsub("\\b[a-zA-Z]\\b{1}", " ", x) #omits singletons including Aa and Ii
  
  x <- tolower(x)
  
  
  #splits character object to data frame, one word per row
  x <- split_string(x)
  
  
  
  #lemmatize string
  
  return(x)
  
}



```




<br>

## New full cleaning function


```{r}


#load in the datasets that are needed to remove stopwords and 
load(url("https://github.com/fsleyes/NLP_S2026/raw/refs/heads/main/week3/utils/replacements_25.rda"))
load(url("https://github.com/fsleyes/NLP_S2026/raw/refs/heads/main/week3/utils/MIT_stops.rda"))


#x is a character object (string), NOT a data frame
shawn_Cleaning_Genie <- function(x) {
  
  
  #start as character object
  x <- tolower(x) #convert to lowercase
  x <- stripWhitespace(x) #strip whitespace
  x <- stri_remove_empty(x) #removes empty strings
  x <- gsub("\\b[a-zA-Z]\\b{1}", " ", x) # omits singletons including Aa and Ii
  x <- split_string(x) #pivots to one word per row
  
  
  #load in the list of contractions that we want to replace
  contractions_list <- replacements_25$word
  
  
  x <- x %>%
    mutate(new_words = case_when( #create new variable called new_words that will have the replaced words
      is.na(string) ~ NA, #if empty then do nothing
      string %in% contractions_list ~ replacements_25$replacement[match(string, replacements_25$word)], #if word is a contraction, replace it with the expanded version
      TRUE ~ string), #if it's none of those options, just keep the original
      check = ifelse(is.na(string), NA, string %in% contractions_list), #this is for debugging
      string = new_words #reassign to the string name
    ) %>%
    select(string) #delete the intermediate variables
  x <- keep_alphab(x) #remove all special characters

  x$string <- lemmatize_strings(x$string) #lemmatize the words, remove affixes etc
  
  
  
  #load in list of stopwords to remove
  stopwords <- tolower(MIT_stops$word)
  x <- x %>%
    mutate(
      is_stopword = ifelse( #create new variable identifying which words are stopwords
        is.na(string), NA, #if empty do nothing
        string %in% stopwords) #if not empty then TRUE if stopword, FALSE if not stopword
      ) %>%
    filter(is_stopword == FALSE) %>% #keep only the words that are NOT stopwords
    select(string) #delete intermediate variables

  
  x$string <- gsub(x = x$string, pattern = "\\d+", replacement = " ") #remove numbers
  
  
  
  x <- split_string(x) #this last step removes all other whitespace and just cleans up in general
  
  return(x)
  
}


```







<br>
<br>
<br>
<br>




# EXAMPLES


## Cleaning the Unabomber Manifesto

Loading in the data, applying the cleaning function, and converting to corpus/document feature matrix:
```{r}

#load in text as a single character vector
unabomb_raw <- paste(readLines('https://raw.githubusercontent.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/main/unabomber_manifesto.txt'))






#run cleaning genie
unabomb_clean_df <- shawn_Cleaning_Genie(unabomb_raw) %>%
  summarize(author = "TK",
            doc_text = paste(string, collapse = " "))

unabomb_corpus <- corpus(unabomb_clean_df, 
                         doc_id_field = "author",
                         text_field = "doc_text")

tokens_unabomb <- tokens(unabomb_corpus)
dfm_unabomb <- dfm(tokens_unabomb)

names(unabomb_corpus) <- "Unabomber Manifesto"
```

<br>

## Getting some basic off-the-shelf statistics about your corpus

dfm_unabomb has to be a document frame matrix type object here:
```{r}
#lexical diversity

print(textstat_lexdiv(dfm_unabomb, measure = c("TTR", "C", "R", "CTTR", "U", "S", "K", "I", "D", "Vm", "Maas"), log.base = 10))

```

<br>

## Table of lexical characteristcs pre/post

```{r}
#summary gives you the types and tokens
summary(unabomb_corpus)


unabomb_raw_df <- data.frame( #creating a new data frame here to give it some info it wants
  author = "TK",
  doc_text = unabomb_raw
) %>%
  summarize(author = unique(author), #group by author - this might change when there are multiple works per author
            doc_text = paste(doc_text, collapse = " ")) #collapse the text all back into one cell for the corpus conversion
unabomb_raw_corpus <- corpus(unabomb_raw_df, doc_id_field = "author", text_field = "doc_text") #convert to corpus
summary(unabomb_raw_corpus) #take a look at type and tokens
```



<br>

## Lexical Dispersion Plot


Pick the words that you want to display on your dispersion plot:
```{r}



textplot_xray(
    kwic(tokens(unabomb_corpus), pattern = "capitalism"),
    kwic(tokens(unabomb_corpus), pattern = "industrial"),
    kwic(tokens(unabomb_corpus), pattern = "socialism"),
    kwic(tokens(unabomb_corpus), pattern = "science"),
    kwic(tokens(unabomb_corpus), pattern = "intelligence"),
    scale = "absolute"
)
```

<br>

## Word Cloud

You can do a word cloud from quanteda:
```{r}
quanteda.textplots::textplot_wordcloud(dfm_unabomb, random_order = FALSE, rotation = .25, color = "red")
```



<br>

## Document Feature Matrix

You can also find the document feature matrix using topfeatures where you give it how many words you want:
```{r}


#print top 20 words
topfeatures(dfm_unabomb, 20)
```