---
title: "NLP CHEATSHEET"
author: "Shawn Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    css: "stylesTemplates/sw_1_professional.css"
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", 
           "RCurl", "psych", "knitr", 
           "RColorBrewer", 'dplyr', "tidytext",
           "tm", "stringi", "textstem", "psych", "ggdist", 
           "viridis", "zoo",  "plotly",
           install=T)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'gutenbergr', 'janeaustenr', 'quanteda', 'readtext', install=T)
library("quanteda.textplots")
library(quanteda.textstats)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)

library(tidytext)


#unabomber text as example
unabomb_raw <- paste(readLines('https://raw.githubusercontent.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/main/unabomber_manifesto.txt'))

```





# DATA WRANGLING


## Functions to add numbers
```{r}
#function to add 2 numbers

add_func <- function(x, y) {
  
  #add them together
  sum <- x + y
  return(sum)
}


#EXAMPLE: 3 + 4
add_func(3, 4)


```

<br>

## Function to mean 2 numbers
```{r}


mean_func <- function(x, y) {
  
  #take the sum
  sum <- x + y
  
  #divide by 2
  mean <- sum / 2
  return(mean)
}


#EXAMPLE: average of 3 and 4

mean_func(3, 4)

```

<br>




# TEXT CLEANING

## Function to remove any word from a string
```{r}

#old_Word is the word you want to replace
#replacement_Word is the word you want to replace it with
#takes a string returns a string
replace_func <- function(string, old_Word, replacement_Word) {
  
  x <- string
  x_clean <- x %>%
    gsub(pattern = old_Word, replacement = replacement_Word) 
  
  return(x_clean)
  
  
}


#EXAMPLE: removing "little" from mary had a little lamb:
test_string <- "mary had a little lamb"
replace_func(string = test_string, old_Word = "little", replacement_Word = "")


```
<br>

## Function to remove special characters
```{r}

#takes a string returns a string
remove_special_func <- function(myString) {
  

  #replaces punctuation and special characters
  x_clean <- gsub("[[:punct:]]", "", myString)

  
  return(x_clean)
  
  
}


#EXAMPLE: removing punctuation marks from messy string:
test_string <- "ma';ry had$ #@a litt#%le lam%^b"
remove_special_func(test_string)


```
<br>

## Function to split string into data frame, one word per row
```{r}

#takes a string returns a data frame
split_string <- function(x) {
  
  #convert to a data frame first
  cleaned_x <- data.frame(string = x) %>%
    separate_longer_delim(string, delim = " ") %>% #pivots to long format based on delimeter of space
    filter(string != "") #filters so that empty rows are deleted
  return(cleaned_x)
  
}


#EXAMPLE: splitting and unlisting "mary had a little lamb"
test_string <- "mary had a little lamb"
split_string(test_string)



```




<br>

## Remove numbers
```{r}


#takes a string returns a string
remove_numbers <- function(x) {
  
  #uses regex to remove any numbers from a string
  x_clean <- x %>%
    gsub(pattern = "\\d+", replacement = "")
  return(x_clean)
  
}


#EXAMPLE: removing numbers from messy string
test_string <- "mar56854864y h56ad3412 a346 li346tt34634le lam346b"
remove_numbers(test_string)

```

<br>

## Remove non-alphabetic characters, numbers, punctuation
```{r}
keep_alphab <- function(x) {
  
  #uses regex to remove any numbers from a string
  x_clean <- x %>%
    mutate(string = gsub(pattern = "[^a-zA-Z]", replacement = "", string))
    
    
  
  return(x_clean)
  
}


#EXAMPLE: removing all these from messy string as a split and unlisted data frame
test_string <- "ma245ry[][]][] ha]][d a 78990lit35567%#$tl87e[][] lam36;;'[][537b"
test_string_split <- split_string(test_string)
keep_alphab(test_string_split)
```



<br>






<br>

## Full cleaning genie

Note: This requires previous functions keep_alphab() and split_string()
```{r}


#x is a character object (string), NOT a data frame
shawn_genie <- function(x) {
  
  
  #start as character object
  x <- tolower(x) #convert to lowercase
  x <- stripWhitespace(x) #strip whitespace
  x <- stri_remove_empty(x) #removes empty strings
  x <- gsub("\\b[a-zA-Z]\\b{1}", " ", x) # omits singletons including Aa and Ii
  x <- split_string(x) #pivots to one word per row
  
  
  load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/replacements_25.rda?raw=true"))
  
  #load in the list of contractions that we want to replace
  contractions_list <- replacements_25$word
  
  
  x <- x %>%
    mutate(new_words = case_when( #create new variable called new_words that will have the replaced words
      is.na(string) ~ NA, #if empty then do nothing
      string %in% contractions_list ~ replacements_25$replacement[match(string, replacements_25$word)], #if word is a contraction, replace it with the expanded version
      TRUE ~ string), #if it's none of those options, just keep the original
      check = ifelse(is.na(string), NA, string %in% contractions_list), #this is for debugging
      string = new_words #reassign to the string name
    ) %>%
    select(string) #delete the intermediate variables
  x <- keep_alphab(x) #remove all special characters

  x$string <- lemmatize_strings(x$string) #lemmatize the words, remove affixes etc
  
  
  load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Temple_stops25.rda?raw=true"))
  
  #load in list of stopwords to remove
  stopwords <- tolower(Temple_stops25$word)
  x <- x %>%
    mutate(
      is_stopword = ifelse( #create new variable identifying which words are stopwords
        is.na(string), NA, #if empty do nothing
        string %in% stopwords) #if not empty then TRUE if stopword, FALSE if not stopword
      ) %>%
    filter(is_stopword == FALSE) %>% #keep only the words that are NOT stopwords
    select(string) #delete intermediate variables

  
  x$string <- gsub(x = x$string, pattern = "\\d+", replacement = " ") #remove numbers
  
  
  
  x <- split_string(x) #this last step removes all other whitespace and just cleans up in general
  
  return(x)
  
}



# EXAMPLE:


#run cleaning genie
unabomb_clean_df <- shawn_genie(unabomb_raw) %>%
  dplyr::summarize(author = "TK",
            doc_text = paste(string, collapse = " "))

unabomb_clean <- shawn_genie(unabomb_raw)





```



## Convert into corpus, tokenize, and turn into document feature matrix



```{r}
#corpus() is used to turn a text data frame with one row per document, columns for metadata then one column for text
unabomb_corpus <- corpus(unabomb_clean_df, 
                         doc_id_field = "author",
                         text_field = "doc_text")

#tokenize
tokens_unabomb <- tokens(unabomb_corpus)

#turn the tokenized object into document feature matrix
dfm_unabomb <- dfm(tokens_unabomb)

names(unabomb_corpus) <- "Unabomber Manifesto"

```





<br>
<br>

# CORPUS STATISTICS


## Getting some basic off-the-shelf statistics about your corpus

dfm_unabomb has to be a document frame matrix type object here:
```{r}
#lexical diversity, these are all different measures
print(textstat_lexdiv(dfm_unabomb, measure = c("TTR", "C", "R", "CTTR", "U", "S", "K", "I", "D", "Vm", "Maas"), log.base = 10))

```



<br>


## Table of lexical characteristcs pre/post

```{r}
#summary gives you the types and tokens
summary(unabomb_corpus)


unabomb_raw_df <- data.frame( #creating a new data frame here to give it some info it wants
  author = "TK",
  doc_text = unabomb_raw
) %>%
  summarize(author = unique(author), #group by author - this might change when there are multiple works per author
            doc_text = paste(doc_text, collapse = " ")) #collapse the text all back into one cell for the corpus conversion
unabomb_raw_corpus <- corpus(unabomb_raw_df, doc_id_field = "author", text_field = "doc_text") #convert to corpus
summary(unabomb_raw_corpus) #take a look at type and tokens
```

<br>


## Document Feature Matrix

You can also find the document feature matrix using topfeatures where you give it how many words you want:
```{r}


#print top 20 words
topfeatures(dfm_unabomb, 20)
```


<br>


# SENTIMENT ANALYSIS


## Merge with lookup tables:

We merge with a lookup table to assign indices to each word that exists in our corpus.
```{r}

#load in database from online
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/lookup_Jul25.rda?raw=true"))

#cut the database to the variables that we're interested in
lookup_small <- lookup_Jul25 %>% select(word, emo_arousal_b24, emo_intensity, emo_trust, emo_valence_b24, lex_AoA)

df_prep <- unabomb_clean %>% left_join(lookup_small, by = c("string" = "word")) 
```


## Descriptive statistics of sentiment indices

describe() from psych package is an easy way of getting basic descriptive statistics from columns extracted from databases.
```{r}
df_prep %>%
  select(-string) %>%
  psych::describe() %>%
  select(mean, sd, median)
  
```






# CORPUS PLOTS



## Lexical Dispersion Plot


Pick the words that you want to display on your dispersion plot:
```{r}

textplot_xray(
    kwic(tokens(unabomb_corpus), pattern = "capitalism"),
    kwic(tokens(unabomb_corpus), pattern = "industrial"),
    kwic(tokens(unabomb_corpus), pattern = "socialism"),
    kwic(tokens(unabomb_corpus), pattern = "science"),
    kwic(tokens(unabomb_corpus), pattern = "intelligence"),
    scale = "absolute"
)
```


<br>

## Word Cloud

You can do a word cloud from quanteda:
```{r}
quanteda.textplots::textplot_wordcloud(dfm_unabomb, random_order = FALSE, rotation = .25, color = "red")
```


<br>



## Histograms

Graphical display of a specific variable of interest in terms of its representation in the corpus, as a histogram.
```{r}
df_prep %>%
  ggplot(aes(x = emo_arousal_b24, fill = after_stat(x))) +
  geom_histogram(color = "white") +
  scale_fill_gradient(low = "#3498db", high = "#e74c3c") + 
  theme_classic()
```


<br>


## Intensity ratings across the corpus

Tracking a variable of interest over the span of the corpus.
```{r}
df_prep %>% mutate(smooth_intensity = zoo::rollmean(emo_intensity, k = 5, fill = NA),
                             word_num = row_number()) %>%
  ggplot(mapping = aes(x = word_num, y = smooth_intensity)) +
  geom_col(position = "dodge") +
  theme_classic()
```


<br>


## Correlation plot

Correlation plots can show general trends of relationships between variables:

```{r}
library(Hmisc)
library(reshape)

cor_df <- df_prep %>%
  select(-string)

get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
}

cormat <- rcorr(as.matrix(cor_df))
cormat_r <- cormat$r
cormat_r_upper <- get_upper_tri(cormat_r)
cormat_r_melt <- melt(cormat_r_upper, na.rm = TRUE)
names(cormat_r_melt)[3] <- "correlation"


cormat_r_melt %>%
  ggplot(mapping = aes(x = X1, y = X2, fill = correlation)) +
  geom_tile(color = "black") + 
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
   midpoint = 0, limit = c(-1, 1), space = "Lab", 
   name="Correlation") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 5, hjust = 1),
    axis.text.y = element_text(size = 5))+
 coord_fixed()

```


<br>


## Line graphs

These  line graphs track a variable of interest over the course of the corpus in a linear representation.

```{r}
df_prep %>%
  mutate(word_num = row_number()) %>%
  ggplot(mapping = aes(x = word_num, y = lex_AoA)) +
  geom_smooth(se = FALSE, 
              aes(color = after_stat(y))) +
  scale_color_gradient(low = "blue", high = "red") + 
  theme_classic()
```
