---
title: "NLP CHEATSHEET"
author: "Shawn Wang"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    css: !expr here::here("./stylesTemplates/sw_1.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)

library(tidytext)

```


# FUNCTIONS



## Functions to add numbers
```{r}
#function to add 2 numbers

add_func <- function(x, y) {
  
  #add them together
  sum <- x + y
  return(sum)
}



```



## Function to mean 2 numbers
```{r}


mean_func <- function(x, y) {
  
  #take the sum
  sum <- x + y
  
  #divide by 2
  mean <- sum / 2
  return(mean)
}


```



## Function to remove any word from a string
```{r}

#old_Word is the word you want to replace
#replacement_Word is the word you want to replace it with
#takes a string returns a string
replace_func <- function(string, old_Word, replacement_Word) {
  
  x <- string
  x_clean <- x %>%
    gsub(pattern = old_Word, replacement = replacement_Word) 
  
  return(x_clean)
  
  
}


```


## Function to remove special characters
```{r}

#takes a string returns a string
remove_special_func <- function(myString) {
  

  #replaces punctuation and special characters with blanks
  x_clean <- gsub("[[:punct:]]", " ", myString)

  
  return(x_clean)
  
  
}


```


## Function to split string into data frame, one word per row
```{r}

#takes a string returns a data frame
split_string <- function(x) {
  
  #convert to a data frame first
  cleaned_x <- data.frame(string = x) %>%
    separate_longer_delim(string, delim = " ") %>% #pivots to long format based on delimeter of space
    filter(string != "") #filters so that empty rows are deleted
  return(cleaned_x)
  
}



```
## Function to remove "the"
```{r}


remove_the <- function(x) {
  
  #replaces "the" with blank
  x_clean <- x %>%
    gsub(pattern = "the", replacement = "", ignore.case = TRUE)
  return(x_clean)
}

```



## Remove numbers
```{r}


#takes a string returns a string
remove_numbers <- function(x) {
  
  #uses regex to remove any numbers from a string
  x_clean <- x %>%
    gsub(pattern = "\\d+", replacement = " ")
  return(x_clean)
  
}

```


## Remove non-alphabetic characters, numbers, punctuation
```{r}
keep_alphab <- function(x) {
  
  #uses regex to remove any numbers from a string
  x_clean <- x %>%
    mutate(string = gsub(pattern = "[^a-zA-Z]", replacement = " ", string))
    
    
  
  return(x_clean)
  
}
```




## Function to clean built from other functions

```{r}

#takes a string, returns a tidy data frame with one word per row
#x is the data frame
#words is a list of words that we want to remove as a vector
clean_string <- function(x, words) {
  

  
  #keep only alphabetic characters
  #remove non-alphabetic characters, numbers, punctuation
  x <- keep_alphab(x)
  
  
  
  #remove words that we don't want by sequencing along a list of words that we don't want in a loop
  for (i in seq_along(words)) {
    
    x <- replace_func(string = x, old_Word = words[i], replacement_Word = "")
    
  }
  
  x <- gsub("\\b[a-zA-Z]\\b{1}", " ", x) #omits singletons including Aa and Ii
  
  x <- tolower(x)
  
  
  #splits character object to data frame, one word per row
  x <- split_string(x)
  
  
  
  #lemmatize string
  
  return(x)
  
}



```





## New full cleaning function


```{r}

#test <- unabomb
test <- PnP

test <- tolower(test)
test <- stripWhitespace(test)
test <- stri_remove_empty(test)
test <- gsub("\\b[a-zA-Z]\\b{1}", " ", test) # omits singletons including Aa and Ii
test <- split_string(test)

contractions_list <- replacements_25$word
test <- test %>%
  mutate(new_words = case_when(
    is.na(string) ~ NA,
    string %in% contractions_list ~ replacements_25$replacement[match(string, replacements_25$word)],
    TRUE ~ string),
    check = ifelse(is.na(string), NA, string %in% contractions_list),
    string = new_words
  ) %>%
  select(string)
test <- keep_alphab(test)
test <- split_string(test)
test$string <- lemmatize_strings(test$string)


stopwords <- tolower(MIT_stops$word)
test <- test %>%
  mutate(
    is_stopword = ifelse(
      is.na(string), NA,
      string %in% stopwords)
    ) %>%
  filter(is_stopword == FALSE) %>%
  select(string)

test$string <- remove_numbers(test$string)



```