---
title: "Multidocument Corpora"
author: "Jamie Reilly, Ph.D."
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  
library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'quanteda', 'tidyr', 'textstem', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 11) %>%
   row_spec(0, color = "yellow", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "300px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
#read csv data file from github
#step1 <- "https://github.com/Reilly-ConceptsCognitionLab/sandbox_scrapfiles/raw/main/fakedat.csv"
#step2 <- read.csv(step1, header=T)
#read string data using 'here' from a local directory
#to_R <- readLines(here("data", "MyRaw.txt")) ---- using 'here' to read in data
#to_format <- paste(to_R, collapse=" ")
#textfromgittoR <- paste(readLines("https://github.com/reilly-lab/TempleTextMine2020/raw/master/unabomber_manifesto.txt"), collapse = " ")  #from github
```

# Load Lookup Databbase
Your lookup database contains values or 'metadata' for many words.  
```{r}
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/lookup_Jul25.rda?raw=true"))
lookup_db <- lookup_Jul25
colnames(lookup_db)
str(lookup_db)
```

# Load and Inspect Gaslighting Corpus
Gaslighting is a multi-document corpus
```{r}
# read raw csv from my computer prep to post on github private
#gaslight <- read.csv('data/gaslight.csv')

#save as *.rda for github
#save(gaslight, file="data/gaslight.rda")

# Load gaslight corpus 
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/gaslight.rda?raw=true"))

str(gaslight)

gaslight$Gender <- as.factor(gaslight$Gender)
gaslight$ID <-  as.factor(gaslight$ID)
gaslight$Ever_Gaslit <- as.factor(gaslight$Ever_Gaslit)
str(gaslight)

#Make sure your variables are structured the way they should be (hint: factors)
levels(gaslight$Gender)

my_tab <- table(gaslight$Ever_Gaslit, gaslight$Gender)
print(my_tab)
summary(my_tab)
```

# Jamie cleaning genie
Here's a cleaning genie that should catch many of the errors and stray dogs that your original cleaning genies missed.
```{r, eval=T}
jamie_genie <- function(dat, wordcol, omit_stops = TRUE, lemmatize = TRUE) {
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(stringi)
  
  # Input validation
  #need to tell R what column in your dataframe is the text in quotes 'wordcol', this error message checks
  if (!wordcol %in% names(dat)) {
    stop(paste("Column", wordcol, "not found in dataframe"))
  }

#load stopword and replacement lists from github  
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Temple_stops25.rda?raw=true"))
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/replacements_25.rda?raw=true"))


# Create working copy and perform initial split, standardizes text encoding so all characters are UTF8, adds an id variable
# transform all text to lower
  dat_prep <- dat %>% dplyr::mutate(id_row_orig = factor(seq_len(nrow(dat))),
      text_initialsplit = tryCatch(stringi::stri_enc_toutf8(as.character(.[[wordcol]]),
                                 is_unknown_8bit = TRUE,
                                 validate = TRUE), error = function(e) stringi::stri_encode(as.character(.[[wordcol]]), to = "UTF-8")
      ) %>% tolower())

  # Standardize apostrophes in original text
  dat_prep <- dat_prep %>% dplyr::mutate(text_initialsplit = ifelse(
        is.na(text_initialsplit), NA_character_, stringi::stri_replace_all_regex(
          text_initialsplit,
          "[\u2018\u2019\u02BC\u201B\uFF07\u0092\u0091\u0060\u00B4\u2032\u2035]",
          "'"
        )
      )
    )

  # Perform initial split into words (retains contractions)
  dat_prep <- dat_prep %>% tidyr::separate_rows(text_initialsplit, sep = "[[:space:]]+") %>%
    dplyr::mutate(
      text_initialsplit = ifelse(
        is.na(text_initialsplit) | stringi::stri_isempty(text_initialsplit),
        NA_character_,
        text_initialsplit
      )
    ) %>%
    # Remove original column
    dplyr::select(-all_of(wordcol))

  # Initialize cleaning column with actual cleaned text
  dat_prep <- dat_prep %>%
    dplyr::mutate(
      word_clean = text_initialsplit,
      # Remove non-alphabetic characters except apostrophes
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^a-zA-Z']", " ")
      ),
      # Squish whitespace
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringr::str_squish(word_clean)
      ),
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^a-z']", "")
      ),
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^[:alnum:]']", "")
      )
    )

#Replace words function - this finds contractions and replaces them with their expanded forms from a lookup database. 
 replace_words <- function(word_clean, replacement_df) {
 for (i in seq_len(nrow(replacement_df))) {
   text_vector <- gsub(
     pattern = replacement_df$word[i],
     replacement = replacement_df$replacement[i],
     x = word_clean, ignore.case = TRUE)
 }
 return(word_clean)
 }
 
 #run replace words function on dat_prep
dat_prep <- replace_words(dat_prep, replacement_df = replacements_25)
  
  

  # Perform additional splitting after replacements (keeping NAs)
  dat_prep <- dat_prep %>%
    tidyr::separate_rows(word_clean, sep = "[[:space:]]+", convert = TRUE) %>%
    dplyr::mutate(
      word_clean = ifelse(
        is.na(word_clean) | stringi::stri_isempty(word_clean),
        NA_character_,
        word_clean
      ),
      id_row_postsplit = seq_len(dplyr::n())  # Fixed this line
    )

  # Lemmatization if requested (only for non-NA values)
  if (lemmatize) {
    dat_prep <- dat_prep %>%
      dplyr::mutate(
        word_clean = ifelse(
          is.na(word_clean),
          NA_character_,
          textstem::lemmatize_strings(word_clean)
        )
      )
  }

  # Stopword removal if requested (now keeps NAs)
  if (omit_stops) {
    stopwords <- tolower(Temple_stops25$word)
    dat_prep <- dat_prep %>%
      dplyr::mutate(
        is_stopword = ifelse(
          is.na(word_clean),
          NA,
          word_clean %in% stopwords
        ),
        word_clean = ifelse(is_stopword, NA_character_, word_clean)
      ) %>%
      dplyr::select(-is_stopword)  # Remove the is_stopword column
  }

  return(dat_prep)
}

```

# Clean your corpus
Adjust genie as needed
```{r, eval=F}

gaslight_just_story <- gaslight %>% select(ID, gd_adult_story)
str(gaslight_just_story)

mycleandat <- jamie_genie(gaslight_just_story, 'gd_adult_story')
print(mycleandat) 
```

# Yoke values from lookup database
Choose a few variables (e.g., 3) to characterize gaslighting on
```{r}
#Choose some variables from the lookup database

#Check terms for joining on
my_gaslight <- gaslight_clean %>% dplyr::left_join(lookup, by = c("CHANGE_THIS" = "CHANGE_THIS")) 
```

# Inspect joined data
```{r}
head(my_gaslight, n=125)
```

# Summarize data
Mean, min, max, SD for each variable <br>
Remember to group by your document_ID
```{r}
#group dataframe by ID, then summarize your vars
my_gaslight <- my_gaslight %>% group_by(ID) %>% dplyr::summarize_at(c("VAR1", "VAR2"), list(mean=mean, sd=sd, min = min, max = max), na.rm=T) %>% ungroup()
head(my_gaslight, n=50)
```


# Visualize your data
```{r}

```


## Archive
```{r, eval=F}

# ARCHIVE
# Join lookup data to target text by row
df_prep <- test %>% dplyr::left_join(lookup_small, by = c("doc_lemmatized" = "word")) 
print(df_prep)
```

