---
title: "Semantic Distance"
author: "Jamie Reilly, Ph.D."
date: "`r format(Sys.Date(), '%B %Y')`"
html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 1
    css: "https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/jr_20.css?raw=true"
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", 'stringi', 'quanteda', 'readr', 'textclean', 'readtext', "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'text2vec', 'textmineR', 'topicmodels', 'tm', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
```


# Semantic Distance
Semantic distance is a quantitative metric of 'distance' (or similarity) between 2 or more concepts (as words) situated within an n-dimensional semantic space. <br>

Semantic spaces differ in their biological/psychological plausibility. Early embedding models (LSA, Word2Vec, GLoVE) quantify semantic distance heavily in terms of co-occurence between word pairs in natural language corpora.  More modern embeddings use co-occurence as a base step but often pass this information to generate embeddings 'on the fly' based on lexical context. This is how embeddings like ELMO work. An embedding is typically a long vector of numbers characterizing the 'meaning' of a word. Experiential semantic ratings provide another way of measuring pairwise distance. <br>

The most common mathematical aproach to measuring similarity between two vectors is by computing the cosine of the angle between them.  Cosine ranges from -1 to 1.  A cosine of -1 is observed when two vectors are identical. It is often helpful to convert raw cosine similarity to cosine distance. This is a simple transformation. You just add +1 to the raw cosine similarity value and then you have distance bounded from 0 to 2.  O is no distance between two concepts (or words). 2 is the highest possible distance.<br>

# Our R package
[Link to our R package here](https://github.com/Reilly-ConceptsCognitionLab/SemanticDistance) <br>

```{r}
#Load the package
library(SemanticDistance)
```

# Step 1: Identify Your Data Structure
`Semantic distance` can compute semantic relationships in structured and unstructured text. It also does numerous cleaning operations. First, figure out your data structure. Is it? <br>
1) Bag-of-words <br>
2) Word pairs arrayed in columns <br>
3) Monologues (continuous sequential text where order is maintained) <br>
3) Dialogues (conversation transcripts where a speaker is marked) <br>

```{r}
shawnVector <- data.frame(words = c("koala", "panda", "fox", "anteater", "hyrax", "beaver",
                 "aubergine", "cabbage", "potato", "pomegranate", "banana", "squash",
                 "pizza", "fries", "salad", "dog"))


```


## Clean monologue
```{r}
Monologue_Cleaned <- clean_monologue_or_list(dat=shawnVector, wordcol='words', omit_stops=TRUE, lemmatize=TRUE)
knitr::kable(head(Monologue_Cleaned, 16), format = "pipe", digits=2)
```

## Clean dialogue
Arguments to clean_dialogue <br>
-dat your raw dataframe with at least one column of text AND a talker column. 
-wordcol column name (quoted) containing the text you want cleaned.
-whotalks column name (quoted) containing the talker ID (will convert to factor).
-omit_stops omits stopwords, T/F default is TRUE.
-lemmatize transforms raw word to lemmatized form, T/F default is TRUE.

```{r}
Dialogue_Cleaned <- clean_dialogue(dat=Dialogue_Typical, wordcol="text", who_talking = "speaker", omit_stops=TRUE, lemmatize=TRUE)
knitr::kable(head(Dialogue_Cleaned, 12), format = "pipe", digits=2)
```

## Clean word pairs
Pairs of words arrayed in columns

```{r}

shawnPairs <- data.frame(word1 = c("oatmeal", "pancake", "waffle", "bear"),
                         word2 = c("panda", "fox", "beaver", "koala"))
WordPairs_Clean <- clean_paired_cols(dat=shawnPairs, wordcol1='word1', wordcol2='word2', lemmatize=TRUE)
knitr::kable(head(WordPairs_Clean, 6), format = "simple", digits=2) 
```

# Step 2: Run Distance
## Ngram to word (rolling)
```{r}
Ngram2Word_Dists1 <- dist_ngram2word(dat=Monologue_Cleaned, ngram=1) #distance word-to-word
knitr::kable(head(Ngram2Word_Dists1, 8), format = "pipe", digits = 2)
```

# Anchor to word (fixed)
Computes semantic distance from each new word in your sample to the initial block of n-words.
```{r}
Anchored_Dists1 <- dist_anchor(dat=Monologue_Cleaned, anchor_size=8)
knitr::kable(head(Anchored_Dists1, 4), format = "pipe", digits = 2)
```


## Semantic distance turn-to-turn in dialogues

```{r}
DialogueDists <- dist_dialogue(dat=Dialogue_Cleaned, who_talking='speaker')
knitr::kable(head(DialogueDists, 4), format = "pipe", digits = 2)
```

## Word Pair Distances
```{r}
Columns_Dists <- dist_paired_cols(WordPairs_Clean) #only argument is dataframe
knitr::kable(head(Columns_Dists, 6), format = "pipe", digits = 2)
```


# Visualize data
if you want to plot continuous distance, you will need to 'fill in' or impute missing values. Lots of ways to do this. We will use linear interpolation from the `zoo` package.

### cluster dendrogram
Arguments to clean_monologue_or_list
- datinput formatted dataframe. <br> -dist_typequoted argument distance option, default is "embedding", alt isSD15`.
```{r}
MyCleanList <- clean_monologue_or_list(dat=shawnVector, wordcol='words')
mydendro <- wordlist_to_network(MyCleanList, output='network', dist_type='embedding')
```

### igraph network

```{r}
MyCleanList <- clean_monologue_or_list(dat=Unordered_List, wordcol='mytext')
mynetwork <- wordlist_to_network(MyCleanList, output='network', dist_type='embedding')
```


Create a new data frame
Sample 500,000 random word pairs
Histograms for both distributions
Density plots on top of each other
Compute correlation between SD15 and GLoVe embeddings

```{r}
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/lookup_Jul25.rda?raw=true"))

word_pop <- lookup_Jul25$word
word_sample <- data.frame(word1 = sample(x = word_pop, size = 500000, replace = TRUE),
                          word2 = sample(x = word_pop, size = 500000, replace = TRUE))

WordPairs_Clean <- clean_paired_cols(dat=word_sample, wordcol1='word1', wordcol2='word2', lemmatize=TRUE)
knitr::kable(head(WordPairs_Clean, 6), format = "simple", digits=2) 

Columns_Dists <- dist_paired_cols(WordPairs_Clean) #only argument is dataframe
knitr::kable(head(Columns_Dists, 6), format = "pipe", digits = 2)

```





