---
title: "NLP WEEK 2"
author: "Shawn Wang"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    css: !expr here::here("../stylesTemplates/sw_1.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)

library(tidytext)

```


# Functions


## Functions to add numbers
```{r}
#function to add numbers

add_func <- function(x, y) {
  
  sum <- x + y
  return(sum)
}



```



## Function to mean 2 numbers
```{r}
mean_func <- function(x, y) {
  
  sum <- x + y
  mean <- sum / 2
  return(mean)
}


```



## Function to remove any word from a string
```{r}


replace_func <- function(string, old_Word, replacement_Word) {
  
  x <- string
  x_clean <- x %>%
    gsub(pattern = old_Word, replacement = replacement_Word) 
  
  return(x_clean)
  
  
}


```
## Function to remove special characters
```{r}


remove_special_func <- function(myString) {
  

  
  x_clean <- gsub("[[:punct:]]", "", myString)

  
  return(x_clean)
  
  
}


```


## Function to split string into data frame, one word per row
```{r}


split_string <- function(x) {
  
  cleaned_x <- data.frame(string = x) %>%
    separate_longer_delim(string, delim = " ") %>%
    filter(string != "")
  return(cleaned_x)
  
}



```
## Function to remove "the"
```{r}


remove_the <- function(x) {
  
  x_clean <- x %>%
    gsub(pattern = "the", replacement = "", ignore.case = TRUE)
  return(x_clean)
}

```



## Remove numbers
```{r}



remove_numbers <- function(x) {
  
  
  x_clean <- x %>%
    gsub(pattern = "\\d+", replacement = "")
  return(x_clean)
  
}

```



## Function to clean built from other functions

```{r}


clean_string <- function(x, words) {
  
  #remove special characters
  x <- remove_special_func(x)
  
  
  #remove numbers 
  x <- remove_numbers(x)
  
  #remove words that we don't want
  for (i in seq_along(words)) {
    
    x <- replace_func(string = x, old_Word = words[i], replacement_Word = "")
    x <- tolower(x)
  }
  
  return(x)
  
}



```










# Tasks


## Remove special characters and split into rows "mary had a little lamb"
```{r}


test_string <- "ma%ry h*ad a l$itt@le la***)mb and the far*m #"

#use split_string function to remove special characters
result_string <- split_string(remove_special_func(test_string))
result_string




```




## Remove all "the"s from Pride and Prejudice
```{r}

#read in data
data <- readLines(here::here("./data/PridePrejudice.txt"))


#replace the with blanks
PnP_noThes <- replace_func(string = data, old_Word = "the", replacement_Word = "")


#split into data frame
PnP_noThes_split <- split_string(PnP_noThes) %>%
  filter(string != "")



```



## Clean the "sticks_messy.txt" file

1. Clean punctuation
2. Get rid of "and", "the", and "for
3. Split into tidy format
4. Count how many words
5. Find type token ratio


### Clean the text
```{r}
sticks_messy <- readLines(here::here("./data/sticks_messy.txt"))


#these are specific words we want to remove
removed_words <- c("and", "the", "for")


#run the integrated string cleaning function
sticks_messy_clean <- clean_string(sticks_messy, removed_words)

#split into data frame with one word per row
sticks_messy_clean_split <- split_string(sticks_messy_clean)
# test <- data.frame(string = sticks_messy_clean) %>%
#   unnest_tokens(word, string)



```


### Count number of words

```{r}
#count number of words
num_words <- nrow(sticks_messy_clean_split)
test <- sticks_messy_clean_split %>%
  mutate(test_column = 1) 
test
```


After cleaning, there are `r num_words` words in the corpus. 



### Calculate TTR
```{r}

#group by unique words using group_by and summarizing how many of each unique word there are
sticks_Unique_Words <- sticks_messy_clean_split %>% group_by(string) %>%
  summarize(n = n())
#this resulting data frame has only unique words

#divide unique words by total words to obtain TTR
sticks_TTR <- nrow(sticks_Unique_Words) / nrow(sticks_messy_clean_split)

```

There are `r nrow(sticks_Unique_Words)` unique words in the sticks dataset after cleaning and removing select stop words. Computing the TTR by dividing the number of unique words by the number of total words: $`r nrow(sticks_Unique_Words)` / `r nrow(sticks_messy_clean_split)` = `r round(sticks_TTR, 2)`$.