---
title: "Corpus Analytics and Visualization in Quanteda"
author: "Jamie Reilly, Ph.D."
date: "`r format(Sys.Date(), '%B %d, %Y')`"
  toc_depth: 1
    css: !expr here::here("../StylesTemplates/jr_19.css")
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", 
           "RCurl", "psych", "knitr", 
           "RColorBrewer", 'dplyr', "tidytext",
           "tm", "stringi", "textstem",
           install=T)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'gutenbergr', 'janeaustenr', 'quanteda', 'readtext', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
#read csv data file from github
#step1 <- "https://github.com/Reilly-ConceptsCognitionLab/sandbox_scrapfiles/raw/main/fakedat.csv"
#step2 <- read.csv(step1, header=T)
#read string data using 'here' from a local directory
#to_R <- readLines(here("data", "MyRaw.txt")) ---- using 'here' to read in data
#to_format <- paste(to_R, collapse=" ")
#textfromgittoR <- paste(readLines("https://github.com/reilly-lab/TempleTextMine2020/raw/master/unabomber_manifesto.txt"), collapse = " ")  #from github
```

# Load Some Texts
Here's a sample text string we will work from. This has no missing values and no punctuation. 
```{r}
#unabomber manifesto
unabomb <- paste(readLines('https://raw.githubusercontent.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/main/unabomber_manifesto.txt'))

#Moby Dick
#uses readtext package
mobydick <- as.character(readtext("http://www.gutenberg.org/cache/epub/2701/pg2701.txt"))

load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/sample_text_join.rda?raw=true"))
my_simple_sample <- sample_text_join 

#Note this contains meta-data
print(my_simple_sample)
```
# Prep for Analytics
Let's get analytics on this simple text sample. We will use the `Quanteda` package. <br>
https://quanteda.io/articles/quickstart.html
```{r}
#load the quanteda package
#library(quanteda)
#First step in quanteda is to convert the target text into a corpus retaining metadata. If you are doing this from a dataframe, you would tell quanteda the docid column and the text_field column
my_simple_corpus  <- corpus(dat, docid_field = "author", text_field = "doc_text")

#or read in a character vector as a corpus object here's Moby Dick
md <- corpus(mobydick)
```

# Ask `Quanteda` for analytics
First summarize your carpus object
```{r}
#we'll summarize Moby Dick
summary(md)
```

# kwic (keyword in context)
search for a keyword and show linguistic contexts
```{r}
#tokenize the corpus object of Moby Dick 
md_tokenized <- tokens(md)
kwic(md_tokenized, pattern = "harpoon")
```

# tokenizer
Quanteda has a powerful built in tokenizer that breaks a document into tokens and stores each token as a new document
```{r}
md_tokenized <- tokens(md, remove_numbers = TRUE,  remove_punct = TRUE, remove_separators = FALSE)
```

# document feature matrix
onvert the corpus to a document term matrix -- well to be more precise, a document feature matrix because elements that are not words are also included
```{r}
# Create a tokenized version of the corpus md, removing punctuation and stopwords
tokens_md <- tokens(md, remove_punct = TRUE) %>% tokens_remove(stopwords("SMART"))

# Create a document-feature matrix, applying stemming
md_dfm <- dfm(tokens_md, dfm_stem = TRUE)
topfeatures(md_dfm, 15)

```

# Visualize Frequency Trends
## Plot Visualizations
```{r}
#We will load quanteda and quanteda.textplots
library("quanteda.textplots")
```

## Wordcloud
```{r}
quanteda.textplots::textplot_wordcloud(md_dfm_small, random_order = FALSE, rotation = 0.25) #word cloud or another way
topfeatures(md_dfm, 70)
```

## Lexical Dispersion Plot
This will show you where keywords of interest appear throughout the document using an x-ray plot
```{r}
## Moby Dick
names(mobydick) <- "Moby Dick"

#uses library quanteda.textplots
textplot_xray(
    kwic(tokens(mobydick), pattern = "whale"),
    kwic(tokens(mobydick), pattern = "ahab"),
    kwic(tokens(mobydick), pattern = "ishmael"),
    kwic(tokens(mobydick), pattern = "kill"),
    scale = "absolute"
) 
```

## Text Stats
Off-the-shelf measures of lexical diversity in Quanteda and what they mean <br>
```{r}
# corpus analytics
library(quanteda.textstats)

#lexical diversity, can take a dfm or corpus
print(textstat_lexdiv(md, measure = c("TTR", "C", "R", "CTTR", "U", "S", "K", "I", "D", "Vm", "Maas"), log.base = 10))

#readability - takes a character or corpus object
textstat_readability(mobydick, measure = "Flesch", remove_hyphens = TRUE)
```


## Unambomber
```{r}
#TBD
```




# Load a very large lexical database
Inspect and tell me about it.
```{r}
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/lookup_Jul25.rda?raw=true"))
```




## Archive
```{r, eval=F}
#read from my computer, save as rda then post on github public repo
sample_text_join <- read.csv("utilities/sample_join.csv")
save(sample_text_join, file="utilities/sample_text_join.rda")
```


# Import Unabomber
```{r, eval=F}
#Import and inspect the raw unabomber manifesto, inspect
unabomb <- paste(readLines('https://raw.githubusercontent.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/main/unabomber_manifesto.txt'))

#Print first few lines of unabomber manifesto
cat(unabomb[1:50], sep = "\n")
str(unabomb)
```


# Define Cleaning Genie
```{r, eval=F}
clean_genie <- function(dat) {
  x <- data.frame(dat)  #coerce input to dataframe
  
#load stopword and replacement lists from reillylab public repo
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Temple_stops25.rda?raw=true"))
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/replacements_25.rda?raw=true"))
  
x <- tolower(x) #transform all text in target colummn to lowercase

#define word replacement fn using a 'for' loop
replace_words <- function(text_vector, replacement_df) {
 for (i in seq_len(nrow(replacement_df))) {
   text_vector <- gsub(
     pattern = replacement_df$word[i],
     replacement = replacement_df$replacement[i],
     x = text_vector,
     ignore.case = TRUE
   )
 }
 return(text_vector)
}

x <- replace_words(x, replacement_df = replacements_25)
x <- gsub("[^A-Za-z ]"," ",x) #removes any nonalphabetic characters, including numerals and punctuation. 
x <- str_squish(gsub("\\s+", " ", x)) # RemoveWhitespace

#Split using strsplit function by the occurence of a space. 
x <- strsplit(x, split = " ")

#Flatten this list 
x <-  unlist(x, recursive = TRUE, use.names = TRUE)
 
# Stopword removal if requested
x <- tm::removeWords(x, Temple_stops25$word)

#removes empty rows
x <- x %>% tidyr::drop_na() 
return(x)  
}

```


# Test fn
```{r, eval=F}
mycleandat <- avery_clean_genie(unabomb)
print(mycleandat) 
```

