---
title: "week4_homework"
author: "Shawn Wang"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    css: !expr here::here("../stylesTemplates/sw_1.css")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", 
           "RCurl", "psych", "knitr", 
           "RColorBrewer", 'dplyr', "tidytext",
           "tm", "stringi", "textstem",
           install=T)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'gutenbergr', 'janeaustenr', 'quanteda', 'readtext', install=T)


library("quanteda.textplots")
library(quanteda.textstats)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
```






```{r, include = FALSE}


#load in the datasets that are needed to remove stopwords and 
load(url("https://github.com/fsleyes/NLP_S2026/raw/refs/heads/main/week3/utils/replacements_25.rda"))
load(url("https://github.com/fsleyes/NLP_S2026/raw/refs/heads/main/week3/utils/MIT_stops.rda"))


#x is a character object (string), NOT a data frame
shawn_Cleaning_Genie <- function(x) {
  
  
  #start as character object
  x <- tolower(x) #convert to lowercase
  x <- stripWhitespace(x) #strip whitespace
  x <- stri_remove_empty(x) #removes empty strings
  x <- gsub("\\b[a-zA-Z]\\b{1}", " ", x) # omits singletons including Aa and Ii
  x <- split_string(x) #pivots to one word per row
  
  
  #load in the list of contractions that we want to replace
  contractions_list <- replacements_25$word
  
  
  x <- x %>%
    mutate(new_words = case_when( #create new variable called new_words that will have the replaced words
      is.na(string) ~ NA, #if empty then do nothing
      string %in% contractions_list ~ replacements_25$replacement[match(string, replacements_25$word)], #if word is a contraction, replace it with the expanded version
      TRUE ~ string), #if it's none of those options, just keep the original
      check = ifelse(is.na(string), NA, string %in% contractions_list), #this is for debugging
      string = new_words #reassign to the string name
    ) %>%
    select(string) #delete the intermediate variables
  x <- keep_alphab(x) #remove all special characters

  x$string <- lemmatize_strings(x$string) #lemmatize the words, remove affixes etc
  
  
  
  #load in list of stopwords to remove
  stopwords <- tolower(MIT_stops$word)
  x <- x %>%
    mutate(
      is_stopword = ifelse( #create new variable identifying which words are stopwords
        is.na(string), NA, #if empty do nothing
        string %in% stopwords) #if not empty then TRUE if stopword, FALSE if not stopword
      ) %>%
    filter(is_stopword == FALSE) %>% #keep only the words that are NOT stopwords
    select(string) #delete intermediate variables

  
  x$string <- gsub(x = x$string, pattern = "\\d+", replacement = " ") #remove numbers
  
  
  
  x <- split_string(x) #this last step removes all other whitespace and just cleans up in general
  
  return(x)
  
}


```



```{r, include = FALSE}
keep_alphab <- function(x) {
  
  #uses regex to remove any numbers from a string
  x_clean <- x %>%
    mutate(string = gsub(pattern = "[^a-zA-Z]", replacement = " ", string))
    
    
  
  return(x_clean)
  
}
```


```{r, include = FALSE}

#takes a string returns a data frame
split_string <- function(x) {
  
  #convert to a data frame first
  cleaned_x <- data.frame(string = x) %>%
    separate_longer_delim(string, delim = " ") %>% #pivots to long format based on delimeter of space
    filter(string != "") #filters so that empty rows are deleted
  return(cleaned_x)
  
}



```





# Tasks

Homework:

1. Read in Unabomber
2. Output structure - what form is it coming in? str?
3. Run cleaning genie - import cleaned version as a corpus
4. Table of lexical characteristics pre/post cleaning - number of types/tokens etc.
5. Lexical dispersion plot - pick 5 words
6. Word cloud
7. Print DFM - top 20 words


## Read in Unabomber

Here I am reading in the text files:
```{r}

#load in text as a single character vector
unabomb_raw <- paste(readLines('https://raw.githubusercontent.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/main/unabomber_manifesto.txt'))

```

<br>

## Output structure

```{r}
str(unabomb_raw)

```

The Unabomber Manifesto is being imported as a character vector.

<br>

## Run cleaning genie
This cleaning genie will clean for punctuation and special characters, expand contractions, remove stopwords, lemmatize, etc. to return a tidy data frame where each row is a word. However, for conversion into a corpus, we will need to smash all of those rows of words back into a single cell. We then turn that corpus into tokens and then a document feature matrix. 
```{r}

#run cleaning genie
unabomb_clean_df <- shawn_Cleaning_Genie(unabomb_raw) %>% #clean and tidy the raw text
  summarize(author = "TK", 
            doc_text = paste(string, collapse = " ")) #turn back into a corpus-ready form

  
#this cleaned df is going to be one row with column author and another column doc_text, which you feed into corpus()
unabomb_corpus <- corpus(unabomb_clean_df, 
                         doc_id_field = "author",
                         text_field = "doc_text")


#convert the corpus into tokens
tokens_unabomb <- tokens(unabomb_corpus)

#turn the tokens into a document feature matrix
dfm_unabomb <- dfm(tokens_unabomb)

names(unabomb_corpus) <- "Unabomber Manifesto"
```


<br>


## Some off-the-shelf statistics

Here are some lexical indices from the unabomber manifesto from the quanteda package:
```{r}
#lexical diversity
print(textstat_lexdiv(dfm_unabomb, measure = c("TTR", "C", "R", "CTTR", "U", "S", "K", "I", "D", "Vm", "Maas"), log.base = 10))


```



<br>

## Table of lexical characteristcs pre/post


```{r}
summary(unabomb_corpus)

#this unabomb_raw_df is just the raw data turned into a corpus for comparison purposes but not used in further analysis
unabomb_raw_df <- data.frame(
  author = "TK",
  doc_text = unabomb_raw
) %>%
  summarize(author = unique(author),
            doc_text = paste(doc_text, collapse = " "))
unabomb_raw_corpus <- corpus(unabomb_raw_df, doc_id_field = "author", text_field = "doc_text")
summary(unabomb_raw_corpus)
```

There are 2837 unique words and 15648 words total in the cleaned manifesto, resulting in a TTR of `r round(2837/15648, 2)`. There are 4849 unique words and 38997 words total in the uncleaned manifesto, resulting in a TTR of `r round(4849/38997, 2)`.

<br>


## Lexical Dispersion Plot

We can generate a lexical dispersion plot to display the prevalence of words that we are interested in:
```{r}


#kwic is keywords in context, and searches for the words that we specify for it under "pattern"
#textplot_xray plots the dispersion
textplot_xray(
    kwic(tokens(unabomb_corpus), pattern = "capitalism"),
    kwic(tokens(unabomb_corpus), pattern = "industrial"),
    kwic(tokens(unabomb_corpus), pattern = "socialism"),
    kwic(tokens(unabomb_corpus), pattern = "science"),
    kwic(tokens(unabomb_corpus), pattern = "intelligence"),
    scale = "absolute"
)
```



<br>

## Word Cloud

We can also generate a wordcloud straight from the document feature matrix object using quanteda:
```{r}


quanteda.textplots::textplot_wordcloud(dfm_unabomb, random_order = FALSE, rotation = .25, color = "red")
```


<br>

## Document Feature Matrix

Finally, we can show the top 20 most common words in our cleaned dataset by feeding the document feature matrix object into the topfeatures() function while specifying the number of words we want:
```{r}


#print top 20 words
topfeatures(dfm_unabomb, 20)
```

