---
title: "week4_homework"
author: "Shawn Wang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", 
           "RCurl", "psych", "knitr", 
           "RColorBrewer", 'dplyr', "tidytext",
           "tm", "stringi", "textstem",
           install=T)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'gutenbergr', 'janeaustenr', 'quanteda', 'readtext', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
```


# Functions

## Cleaning Genie
```{r}


#load in the datasets that are needed to remove stopwords and 
load(url("https://github.com/fsleyes/NLP_S2026/raw/refs/heads/main/week3/utils/replacements_25.rda"))
load(url("https://github.com/fsleyes/NLP_S2026/raw/refs/heads/main/week3/utils/MIT_stops.rda"))


#x is a character object (string), NOT a data frame
shawn_Cleaning_Genie <- function(x) {
  
  
  #start as character object
  x <- tolower(x) #convert to lowercase
  x <- stripWhitespace(x) #strip whitespace
  x <- stri_remove_empty(x) #removes empty strings
  x <- gsub("\\b[a-zA-Z]\\b{1}", " ", x) # omits singletons including Aa and Ii
  x <- split_string(x) #pivots to one word per row
  
  
  #load in the list of contractions that we want to replace
  contractions_list <- replacements_25$word
  
  
  x <- x %>%
    mutate(new_words = case_when( #create new variable called new_words that will have the replaced words
      is.na(string) ~ NA, #if empty then do nothing
      string %in% contractions_list ~ replacements_25$replacement[match(string, replacements_25$word)], #if word is a contraction, replace it with the expanded version
      TRUE ~ string), #if it's none of those options, just keep the original
      check = ifelse(is.na(string), NA, string %in% contractions_list), #this is for debugging
      string = new_words #reassign to the string name
    ) %>%
    select(string) #delete the intermediate variables
  x <- keep_alphab(x) #remove all special characters

  x$string <- lemmatize_strings(x$string) #lemmatize the words, remove affixes etc
  
  
  
  #load in list of stopwords to remove
  stopwords <- tolower(MIT_stops$word)
  x <- x %>%
    mutate(
      is_stopword = ifelse( #create new variable identifying which words are stopwords
        is.na(string), NA, #if empty do nothing
        string %in% stopwords) #if not empty then TRUE if stopword, FALSE if not stopword
      ) %>%
    filter(is_stopword == FALSE) %>% #keep only the words that are NOT stopwords
    select(string) #delete intermediate variables

  
  x$string <- gsub(x = x$string, pattern = "\\d+", replacement = " ") #remove numbers
  
  
  
  x <- split_string(x) #this last step removes all other whitespace and just cleans up in general
  
  return(x)
  
}


```


## Remove non-alphabetic characters, numbers, punctuation
```{r}
keep_alphab <- function(x) {
  
  #uses regex to remove any numbers from a string
  x_clean <- x %>%
    mutate(string = gsub(pattern = "[^a-zA-Z]", replacement = " ", string))
    
    
  
  return(x_clean)
  
}
```


## Split string
```{r}

#takes a string returns a data frame
split_string <- function(x) {
  
  #convert to a data frame first
  cleaned_x <- data.frame(string = x) %>%
    separate_longer_delim(string, delim = " ") %>% #pivots to long format based on delimeter of space
    filter(string != "") #filters so that empty rows are deleted
  return(cleaned_x)
  
}



```





# Load some files

```{r}
#unabomber manifesto
unabomb <- paste(readLines('https://raw.githubusercontent.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/main/unabomber_manifesto.txt'))

#Moby Dick
#uses readtext package
mobydick <- as.character(readtext("http://www.gutenberg.org/cache/epub/2701/pg2701.txt"))

load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/sample_text_join.rda?raw=true"))
my_simple_sample <- sample_text_join 

#Note this contains meta-data
print(my_simple_sample)
```




# Prepare for analyses

## Convert to corpus using Quanteda package
```{r}

#specify the IDs as well as the text
simple_corpus <- corpus(my_simple_sample, 
                        doc_id_field = "author", 
                        text_field = "doc_text")


#you can also load in a singular text file as a character object as a corpus
mobyDick <- corpus(mobydick)
```



## Summarize corpus
```{r}
summary(mobyDick)
```


## kwic
Keyword in context searches for a keyword and shows linguistic contexts

```{r}
md_tokenized <- tokens(mobyDick)
kwic(md_tokenized, pattern = "salt")
```


## Quanteda's cleaning function

```{r}
md_tokenized <- tokens(mobyDick, remove_numbers = TRUE, remove_punct = TRUE, remove_separators = FALSE)
```



## Document Feature Matrix
Converts corpus to document feature matrix (not a document term matrix) because there are features that are not words that are included currently
```{r}

#remove punctuation and remove stopwords
tokens_md <- tokens(mobyDick, remove_punct = TRUE) %>% tokens_remove(stopwords("SMART"))

md_dfm <- dfm(tokens_md)
topfeatures(md_dfm, 15)
```

# Visualization

```{r}
library("quanteda.textplots")
library(quanteda.textstats)
```


## Word Cloud

```{r}


quanteda.textplots::textplot_wordcloud(md_dfm, random_order = FALSE, rotation = .25, color = "red")

topfeatures(md_dfm, 70)
```



## Plot DFM

```{r}
plot(md_dfm)
```



## Lexical dispertion plot

```{r}
names(mobydick) <- "Moby Dick"

textplot_xray(
    kwic(tokens(mobydick), pattern = "whale"),
    kwic(tokens(mobydick), pattern = "ahab"),
    kwic(tokens(mobydick), pattern = "ishmael"),
    kwic(tokens(mobydick), pattern = "kill"),
    scale = "absolute"
)
```



## Text stats
Off the shelf from Quanteda package
```{r}

#lexical diversity
print(textstat_lexdiv(md_dfm, measure = c("TTR", "C", "R", "CTTR", "U", "S", "K", "I", "D", "Vm", "Maas"), log.base = 10))


#readability
textstat_readability(mobyDick, measure = "Flesch", remove_hyphens = TRUE)
```




Homework:



