---
title: "Database Joins & Sentiment Analysis"
author: "Jamie Reilly, Ph.D."
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align = 'center', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  
library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'quanteda', 'tidyr', 'textstem', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 11) %>%
   row_spec(0, color = "yellow", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "300px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
#read csv data file from github
#step1 <- "https://github.com/Reilly-ConceptsCognitionLab/sandbox_scrapfiles/raw/main/fakedat.csv"
#step2 <- read.csv(step1, header=T)
#read string data using 'here' from a local directory
#to_R <- readLines(here("data", "MyRaw.txt")) ---- using 'here' to read in data
#to_format <- paste(to_R, collapse=" ")
#textfromgittoR <- paste(readLines("https://github.com/reilly-lab/TempleTextMine2020/raw/master/unabomber_manifesto.txt"), collapse = " ")  #from github
```

# Demo Text
Here's a sample text string we will work from. This has no missing values and no punctuation. 
```{r}
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/sample_text_join.rda?raw=true"))
my_join <- sample_text_join
print(my_join)

#split and unlist on doc_text dat, textcol
my_join_tidy <- tidyr::separate_rows(my_join, doc_text, sep=" ") 
print(my_join_tidy)

#lemmatize strings in doc_text write new column of lemmatized strings
test <- my_join_tidy %>% mutate(doc_lemmatized = lemmatize_strings(doc_text))

print(test)

```

# Where to find lookup databases?
https://www.reilly-coglab.com/data


# Load BIG lookup database
Your lookup database contains values or 'metadata' for many words.  I have combined many published lexical databases into a single lexical lookup database that you can use for sentiment, etc.  Let's load that in from Github and inspect it
```{r}
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/lookup_Jul25.rda?raw=true"))
lookup_db <- lookup_Jul25
colnames(lookup_db)
str(lookup_db)
```


# Let's choose five variables
We can do some basic sentiment analysis or any other variable in the lookup database
```{r}
lookup_small <- lookup_db %>% select(1, 2, 24, 27, 28)
str(lookup_small)
```

# Join lookup data to target text by row
```{r}
df_prep <- test %>% dplyr::left_join(lookup_small, by = c("doc_lemmatized" = "word")) 
print(df_prep)
```



# Now do it on unambomber and visualize the results
# Import Unabomber
```{r, eval=T}
#Import and inspect the raw unabomber manifesto, inspect
unabomb <- paste(readLines('https://raw.githubusercontent.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/main/unabomber_manifesto.txt'))

#Print first few lines of unabomber manifesto
cat(unabomb[1:50], sep = "\n")
str(unabomb)
```


# Jamie cleaning genie
Here's a cleaning genie that should catch many of the errors and stray dogs that your original cleaning genies missed.
```{r, eval=F}
jamie_genie <- function(dat, wordcol, omit_stops = TRUE, lemmatize = TRUE) {
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(stringi)
  
  # Input validation
  #need to tell R what column in your dataframe is the text in quotes 'wordcol', this error message checks
  if (!wordcol %in% names(dat)) {
    stop(paste("Column", wordcol, "not found in dataframe"))
  }

#load stopword and replacement lists from github  
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Temple_stops25.rda?raw=true"))
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/replacements_25.rda?raw=true"))


# Create working copy and perform initial split, standardizes text encoding so all characters are UTF8, adds an id variable
# transform all text to lower
  dat_prep <- dat %>% dplyr::mutate(id_row_orig = factor(seq_len(nrow(dat))),
      text_initialsplit = tryCatch(stringi::stri_enc_toutf8(as.character(.[[wordcol]]),
                                 is_unknown_8bit = TRUE,
                                 validate = TRUE), error = function(e) stringi::stri_encode(as.character(.[[wordcol]]), to = "UTF-8")
      ) %>% tolower())

  # Standardize apostrophes in original text
  dat_prep <- dat_prep %>% dplyr::mutate(text_initialsplit = ifelse(
        is.na(text_initialsplit), NA_character_, stringi::stri_replace_all_regex(
          text_initialsplit,
          "[\u2018\u2019\u02BC\u201B\uFF07\u0092\u0091\u0060\u00B4\u2032\u2035]",
          "'"
        )
      )
    )

  # Perform initial split into words (retains contractions)
  dat_prep <- dat_prep %>% tidyr::separate_rows(text_initialsplit, sep = "[[:space:]]+") %>%
    dplyr::mutate(
      text_initialsplit = ifelse(
        is.na(text_initialsplit) | stringi::stri_isempty(text_initialsplit),
        NA_character_,
        text_initialsplit
      )
    ) %>%
    # Remove original column
    dplyr::select(-all_of(wordcol))

  # Initialize cleaning column with actual cleaned text
  dat_prep <- dat_prep %>%
    dplyr::mutate(
      word_clean = text_initialsplit,
      # Remove non-alphabetic characters except apostrophes
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^a-zA-Z']", " ")
      ),
      # Squish whitespace
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringr::str_squish(word_clean)
      ),
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^a-z']", "")
      ),
      word_clean = ifelse(
        is.na(word_clean),
        NA_character_,
        stringi::stri_replace_all_regex(word_clean, "[^[:alnum:]']", "")
      )
    )

  # Apply contractions replacement (only for non-NA values)
  dat_prep <- replacements_25(dat = dat_prep, wordcol = word_clean)

  # Perform additional splitting after replacements (keeping NAs)
  dat_prep <- dat_prep %>%
    tidyr::separate_rows(word_clean, sep = "[[:space:]]+", convert = TRUE) %>%
    dplyr::mutate(
      word_clean = ifelse(
        is.na(word_clean) | stringi::stri_isempty(word_clean),
        NA_character_,
        word_clean
      ),
      id_row_postsplit = seq_len(dplyr::n())  # Fixed this line
    )

  # Lemmatization if requested (only for non-NA values)
  if (lemmatize) {
    dat_prep <- dat_prep %>%
      dplyr::mutate(
        word_clean = ifelse(
          is.na(word_clean),
          NA_character_,
          textstem::lemmatize_strings(word_clean)
        )
      )
  }

  # Stopword removal if requested (now keeps NAs)
  if (omit_stops) {
    stopwords <- tolower(Temple_stops25$word)
    dat_prep <- dat_prep %>%
      dplyr::mutate(
        is_stopword = ifelse(
          is.na(word_clean),
          NA,
          word_clean %in% stopwords
        ),
        word_clean = ifelse(is_stopword, NA_character_, word_clean)
      ) %>%
      dplyr::select(-is_stopword)  # Remove the is_stopword column
  }

  return(dat_prep)
}

```

# Test clean fn
```{r, eval=F}
mycleandat <- clean_genie(unabomb)
print(mycleandat) 
```



## Archive
```{r, eval=F}
#read from my computer, save as rda then post on github public repo
sample_text_join <- read.csv("utilities/sample_join.csv")
save(sample_text_join, file="utilities/sample_text_join.rda")
```
